A mixture of hidden Markov models to predict the lymphatic spread in head and neck cancer depending on primary tumor location

### 2025-06-02 manuscript.qmd
added first risk discussion

We first analyze the risk predictions for the case where we diagnosed

>>>>

We first analyze the risk predictions for the case where we clinically diagnose no positive LNLs

::: {#fig-oral_cavity_early_no_pos layout-ncol=2}

![](figures/plot_C02_early_no_positive_diagnoses.png){#fig-oral_cavity_no_pos_C02_early}

![](figures/plot_C03_early_no_positive_diagnoses.png){#fig-oral_cavity_no_pos_C03_early}

![](figures/plot_C04_early_no_positive_diagnoses.png){#fig-oral_cavity_no_pos_C04_early}

![](figures/plot_C05_early_no_positive_diagnoses.png){#fig-oral_cavity_no_pos_C05_early}

![](figures/plot_C06_early_no_positive_diagnoses.png){#fig-oral_cavity_no_pos_C06_early}

Probability of occult metastases in LNLs I, II, III, IV and V for early t-stages in oral cavity subsites C02 (tongue), C03 (gum), C04 (floor of mouth), C05 (palate) and C06 (other parts of mouth). The red line marks the 5% threshold for high risk of involvement.
:::

In @fig-oral_cavity_early_no_pos we can see that the model predicts different risks of involvement for each subsite. For subsite C03 (gum), predicts a higher risk of involvement for LNL I. In subsites C02 (tongue) and C05 (palate), the model predicts an above threshold risk of involvement for LNL II. While for subsites C02 (tongue) and C06 (other parts of mouth) all LNLs stay below the 5% threshold. Thus, the model would propose three different treatment plans for these subsites, with C03 (gum) LNL I, C02 (tongue) LNL II and C05 (palate) LNL II being irradiated, while C04 (floor of mouth) and C06 (other parts of mouth) would not need elective irradiation at all, which is a major deviation from the clinical guidelines which irradiate LNLs I, II and III for all oral cavity subsites.
